---
title: "ebpmf_log with smooth factor, an example"
author: "DongyueXie"
date: "2023-09-24"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

## Introduction

We give an example of fitting `ebpmf_log` with smoothed factor, using wavelet smoothing.

## Generate some data

```{r}
set.seed(12345)
n <- 50
p <- 256
K <- 3
snr <- 1

## Step 1: sample U, an orthogonal matrix
rand_semdef_sym_mat <- crossprod(matrix(runif(n * n), n, n))
rand_ortho_mat <- eigen(rand_semdef_sym_mat)$vector[, 1:K]
u_1 <- rand_ortho_mat[, 1]
u_2 <- rand_ortho_mat[, 2]
u_3 <- rand_ortho_mat[, 3]

f1 = c(rep(0,p/8), rep(1, p/4), rep(0, p/4), rep(0, p/4),rep(0,p/8))
f2 = c(rep(0,p/8), rep(0, p/4), rep(1, p/4), rep(0, p/4),rep(0,p/8))
f3 = c(rep(0,p/8), rep(0, p/4), rep(0, p/4), rep(1, p/4),rep(0,p/8))

L = cbind(u_1,u_2,u_3)
FF=cbind(f1,f2,f3)

M = n / 3 * u_1 %*% t(f1) +
  n / 5 * u_2 %*% t(f2) +
  n / 6 * u_3 %*% t(f3)
v = var(c(M))/snr
X = M + matrix(rnorm(n*p,0,sqrt(v)),nrow=n,ncol=p)
Y = matrix(rpois(n*p,exp(X)),nrow=n,ncol = p)
```

## Fit model with ebnm_dwt_haar

We use Haar wavelet DWT for smoothing. 

If the number of columns of the matrix is not a power of 2, or the var_type is 'by_col', the ELBO might not be non-decreasing.

```{r}
library(ebpmf.alpha)
fit = ebpmf_log(Y,flash_control = list(ebnm.fn=c(ebnm::ebnm_point_normal,ebnm_dwt_haar)),var_type = 'constant')
plot(fit$elbo_trace)
# plot factors
for(k in 3:fit$fit_flash$n_factors){
  plot(fit$fit_flash$F_pm[,k],type='l')
}
fit$elbo
```

```{r}
fit = ebpmf_log(Y,flash_control = list(ebnm.fn=c(ebnm::ebnm_point_normal,ebnm_dwt_haar)),var_type = 'by_row')
plot(fit$elbo_trace)
# plot factors
for(k in 3:fit$fit_flash$n_factors){
  plot(fit$fit_flash$F_pm[,k],type='l')
}
fit$elbo
```

```{r}
fit = ebpmf_log(Y,flash_control = list(ebnm.fn=c(ebnm::ebnm_point_normal,ebnm_dwt_haar)),var_type = 'by_col')
plot(fit$elbo_trace)
# plot factors
for(k in 3:fit$fit_flash$n_factors){
  plot(fit$fit_flash$F_pm[,k],type='l')
}
fit$elbo
```

## Fit model with ebnm_ndwt

We use Haar wavelet NDWT for smoothing.

The ELBO might not be non-decreasing.

```{r}
fit = ebpmf_log(Y,flash_control = list(ebnm.fn=c(ebnm::ebnm_point_normal,ebnm_ndwt)),var_type = 'constant')
plot(fit$elbo_trace)
# plot factors
for(k in 3:fit$fit_flash$n_factors){
  plot(fit$fit_flash$F_pm[,k],type='l')
}
fit$elbo
```

```{r}
fit = ebpmf_log(Y,flash_control = list(ebnm.fn=c(ebnm::ebnm_point_normal,ebnm_ndwt)),var_type = 'by_row')
plot(fit$elbo_trace)
# plot factors
for(k in 3:fit$fit_flash$n_factors){
  plot(fit$fit_flash$F_pm[,k],type='l')
}
fit$elbo
```

```{r}
fit = ebpmf_log(Y,flash_control = list(ebnm.fn=c(ebnm::ebnm_point_normal,ebnm_ndwt)),var_type = 'by_col')
plot(fit$elbo_trace)
# plot factors
for(k in 3:fit$fit_flash$n_factors){
  plot(fit$fit_flash$F_pm[,k],type='l')
}
fit$elbo
```
